---
title: "wk2 assignment"
author: "Frederick Jones"
date: "2023-09-11"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the data of survey

The data is in excel sheet and the data is to be loaded from the sheet. 

```{r Load Movie_Ratings data}
data <- read.csv("https://raw.githubusercontent.com/jewelercart/Data606_2023/main/Movie_Ratings%20.csv")
data
```

## Saving data to MySql
```{r}
database_name ='student_records'
user_name = 'jewelercart'
password = "567juhmm_"
```


```{r}
#install.packages("RMySQL")
library(RMySQL)
con <- dbConnect(RMySQL::MySQL(), host = "localhost", dbname = database_name, user = user_name, password=password)
dbSendQuery(con, "SET GLOBAL local_infile = true;") 
dbWriteTable(con, "Movie_Ratings", data, overwrite = TRUE)
dbDisconnect(con)
```
## Getting data from database from server running on local host. 

```{r}

con <- dbConnect(RMySQL::MySQL(), host = "localhost", dbname = database_name, user = user_name, password=password)
df=dbReadTable(con, 'Movie_Ratings')
dbDisconnect(con)
df
```

## Saving data on a server and retreiving data from a server is something that every data scientist do in his career. 
In this assignment, the practice of handling data to and fro a server was achieved. However, the server was running on local host. But the same procedure might be followed to get data from a remote server.
